---
title: "echoverse modules"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{echoverse modules}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include=TRUE}
library(echolocatoR)
```

# Overview

`echolocatoR` orchestrates a suite of modular R packages collectively
called the **echoverse**. While `echolocatoR::finemap_loci()` runs the
full pipeline automatically, each module can also be used independently
for custom workflows.

This vignette provides a quick reference for when to use each module.

# Module reference

## Level 1: Base packages

These provide foundational utilities used by all other modules.

### `echodata`

**Purpose**: Data management, example datasets, column standardization.

**Use independently when you want to**:

- Load bundled fine-mapping results: `BST1`, `LRRK2`, `MEX3C`
- Download full GWAS summary stats: `get_Nalls2019()`, `get_Kunkle2019()`
- Standardize column names across different GWAS formats
- Access the Fine-Mapping Portal: `portal_query()`

```{r echodata-example}
## Bundled fine-mapping results (no download needed)
dat <- echodata::BST1
cat("BST1 locus:", nrow(dat), "SNPs,", ncol(dat), "columns\n")
```

### `echogithub`

**Purpose**: GitHub API utilities for downloading files and releases.

**Use independently when you want to**:

- Download files from GitHub repos programmatically
- Access GitHub release assets

### `downloadR`

**Purpose**: File download helpers with caching.

**Use independently when you want to**:

- Download and cache large files with progress tracking
- Upload data to Zenodo

### `devoptera`

**Purpose**: Developer utilities for the echoverse packages.

## Level 2: Mid-level utilities

### `echotabix`

**Purpose**: Tabix-indexed file querying and liftover.

**Use independently when you want to**:

- Query tabix-indexed VCF or summary stats files by genomic region
- Perform genome build liftover (hg19/hg38)
- Convert between file formats

### `echoLD`

**Purpose**: LD (linkage disequilibrium) matrix computation.

**Use independently when you want to**:

- Compute LD matrices from 1000 Genomes Phase 3, UK Biobank, or custom VCFs
- Filter variants by LD or MAF thresholds

```{r echoLD-example}
## Bundled LD matrix
ld <- echodata::BST1_LD_matrix
cat("LD matrix:", nrow(ld), "x", ncol(ld), "SNPs\n")
```

### `echoconda`

**Purpose**: Conda environment management for external tools.

**Use independently when you want to**:

- Set up conda environments for FINEMAP, PolyFun, or other command-line tools
- Manage Python dependencies from R via basilisk

## Level 3: Analysis modules

### `echofinemap`

**Purpose**: Statistical fine-mapping methods.

**Use independently when you want to**:

- Run individual fine-mapping methods (ABF, SUSIE, FINEMAP, PolyFun+SUSIE)
- Compare results across methods
- See available methods: `echofinemap::lfm()`

```{r echofinemap-example}
## List available fine-mapping methods
echofinemap::lfm()
```

### `echoannot`

**Purpose**: Functional annotation and enrichment.

**Use independently when you want to**:

- Overlap fine-mapped SNPs with ROADMAP epigenomic annotations
- Run GoShifter enrichment tests
- Annotate with NOTT2019/CORCES2020 single-cell data
- Create super summary plots

### `echoplot`

**Purpose**: Locus visualization.

**Use independently when you want to**:

- Create multi-track locus zoom plots
- Visualize fine-mapping posterior probabilities alongside GWAS p-values
- Overlay LD, gene tracks, and annotations

### `catalogueR`

**Purpose**: QTL colocalization via the eQTL Catalogue.

**Use independently when you want to**:

- Query the eQTL Catalogue for colocalization analysis
- Access tissue/cell-type-specific QTL datasets

## Level 4: Pipeline orchestrator

### `echolocatoR`

**Purpose**: Unified fine-mapping pipeline.

This is the package you're using now. It calls all modules above in the
correct order via `finemap_loci()` / `finemap_locus()`.

# Choosing your approach

| Scenario | Recommended approach |
|----------|---------------------|
| First-time fine-mapping | `echolocatoR::finemap_loci()` — handles everything |
| Custom fine-mapping workflow | Use `echofinemap` directly with your own LD matrices |
| Just need LD matrices | Use `echoLD::get_LD()` |
| Just need annotations | Use `echoannot` functions directly |
| Just need plots | Use `echoplot::plot_locus()` |
| Querying large summary stats | Use `echotabix::query()` |
| Post-hoc analysis of results | Use `echodata` to load + `echoannot` to annotate |

# Session info

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>
