---
title: "Get started"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Get started}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include=TRUE}
library(echolocatoR)

has_internet <- tryCatch(
    !is.null(curl::nslookup("github.com", error = FALSE)),
    error = function(e) FALSE
)
```

# Full pipeline

All examples below use data from the Parkinson's disease GWAS by Nalls et al. (2019).

## Prepare `top_SNPs` data.frame

* To enable rapid fine-mapping of many loci, you can create a `top_SNPs` data.frame
which contains the position of the lead/index SNP within each locus.
* `finemap_loci()` (see next step) will then use this info to extract subsets of the
full GWAS/QTL summary statistics using windows centered on each lead/index SNP.
* The `topSS` argument can either be a data.frame, or a path to a topSS file saved somewhere.
Most common tabular data formats (e.g. .tsv, .csv, .xlsx) are accepted.


```{r Prepare top_SNPs data.frame, eval=has_internet}
#### Load example top SNPs (pre-formatted) ####
topSS <- echodata::topSNPs_Nalls2019_raw
#### construct a column mapping object ####
colmap <- echodata::construct_colmap(P = "P, all studies",
                                     Effect = "Beta, all studies",
                                     Locus = "Nearest Gene",
                                     Gene = "QTL Nominated Gene (nearest QTL)")
#### Import top SNPs ####
topSNPs <- echodata::import_topSNPs(
    topSS = echodata::topSNPs_Nalls2019_raw,
    colmap = colmap,
    grouping_vars = "Locus Number")
head(topSNPs)
```

## Path to full summary stats file

* Since a full GWAS summary stats file would be too large to include within *echolocatoR*,
we instead provide an example subset of the full summary stats.

* To simulate how you'd actually use your own full summary stats file, we will save our example dataset to your computer (you can change the path to wherever you like).

* We highly recommend munging your full summary stats using the Bioconductor package [`MungeSumstats`](https://github.com/neurogenomics/MungeSumstats) first. It's easy to use and very robust. It also means you don't have to provide most column mapping arguments in `finemap_loci` when `munged=TRUE`.

Here's an example of how to munge your full summary stats file:

```
fullSS_path <- echodata::example_fullSS(munged = FALSE)
fullSS_path <- MungeSumstats::format_sumstats(path = fullSS_path, ref_genome = "GRCH37")
```

We have already munged the following example summary stats for you.

```{r fullSS, eval=has_internet}
fullSS_path <- echodata::example_fullSS(dataset = "Nalls2019")
```

## Run fine-mapping pipeline

For a full description of all arguments, see `?finemap_loci`.

Here are some key arguments:

* *results_dir*: Where you want to store all of your results.
* *finemap_methods*: Which fine-mapping methods you want to run. For a full list of currently supported methods, run the function `echofinemap::lfm()`.
* *bp_distance*: Controls window size. Specifically, `bp_distance` is the number of basepairs upstream/downstream you want to extract for each locus. For example, if you want a 2Mb window (+/- 1Mb from the lead/index SNP in `top_SNPs`), set `bp_distance=1e+06`.
* *plot_zoom*: Zoom in/out from the center of each locus when producing the multiview plot.
You can adjust this separately from `bp_distance` so that you don't have rerun the whole pipeline each time (locus subsets, LD matrices, and fine-mapping results are all automatically saved in locus-specific folders).

**Note**: Please use the full absolute paths (instead of relative paths) wherever possible (e.g. `results_dir`). This is especially important for the tool *FINEMAP*.

The following call fine-maps two loci (BST1 and MEX3C) using three statistical
fine-mapping methods. Depending on your machine and internet speed, this may
take **10-30 minutes per locus**.

```{r Run fine-mapping pipeline, eval=FALSE}
results <- echolocatoR::finemap_loci(
 fullSS_path = fullSS_path,
 topSNPs = topSNPs,
 loci = c("BST1","MEX3C"),
 LD_reference = "1KGphase3",
 dataset_name = "Nalls23andMe_2019",
 fullSS_genome_build = "hg19",
 bp_distance = 1000,
 finemap_methods = c("ABF","SUSIE","FINEMAP"),
 munged = TRUE)
```

## Inspect results

The output of `finemap_loci()` is a named list of data.tables, one per locus.
Each data.table contains the original GWAS summary statistics augmented with
fine-mapping posterior probabilities and credible set assignments.

Here we use the bundled BST1 example results to show what the output looks like:

```{r inspect-results}
dat <- echodata::BST1
## Top fine-mapped SNPs (consensus across methods)
consensus <- subset(dat, Consensus_SNP == TRUE)
knitr::kable(
    consensus[, c("SNP","CHR","POS","P","mean.PP","Support","Consensus_SNP")],
    caption = "Consensus fine-mapped SNPs for the BST1 locus",
    row.names = FALSE
)
```

# Next steps

- Explore and interpret results: `vignette("explore_results")`
- Visualize loci: `vignette("plotting")`
- Summarise across loci: `vignette("summarise")`
- Fine-map QTL data: `vignette("QTLs")`
- Learn about sub-packages: `vignette("echoverse_modules")`

<hr>

# Session info

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>

<br>
