---
title: "Exploring fine-mapping results"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Exploring fine-mapping results}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include=TRUE}
library(echolocatoR)
```

# Overview

This vignette demonstrates how to explore and interpret the fine-mapping
results produced by `finemap_loci()`. All examples use bundled data
included in `echodata`, so **no internet connection is required**.

# Load example results

`echodata` ships with pre-computed fine-mapping results for three
Parkinson's disease loci from Nalls et al. (2019):

```{r load-data}
dat <- echodata::BST1
dim(dat)
```

Each row is a SNP within the locus window. Key columns include:

- **GWAS columns**: `SNP`, `CHR`, `POS`, `P`, `Effect`, `StdErr`, `MAF`
- **Fine-mapping posterior probabilities**: `ABF.PP`, `SUSIE.PP`, `FINEMAP.PP`
- **Credible set assignments**: `ABF.CS`, `SUSIE.CS`, `FINEMAP.CS`
- **Consensus columns**: `Support` (number of methods placing SNP in a CS),
  `Consensus_SNP`, `mean.PP`, `mean.CS`

```{r show-columns}
## Fine-mapping specific columns
fm_cols <- c("SNP","P","ABF.PP","SUSIE.PP","FINEMAP.PP",
             "Support","Consensus_SNP","mean.PP")
knitr::kable(head(dat[order(dat$mean.PP, decreasing = TRUE), ..fm_cols], 10),
             digits = 4, row.names = FALSE,
             caption = "Top 10 SNPs by mean posterior probability")
```

# Consensus SNPs

Consensus SNPs are those placed in a credible set by multiple fine-mapping
methods. The `Support` column indicates how many methods agree:

```{r consensus}
consensus <- subset(dat, Consensus_SNP == TRUE)
knitr::kable(
    consensus[, c("SNP","POS","P","ABF.PP","SUSIE.PP","FINEMAP.PP",
                   "Support","mean.PP")],
    digits = 4, row.names = FALSE,
    caption = "Consensus fine-mapped SNPs (BST1 locus)"
)
```

# SNP group counts

Summarise how many SNPs fall into each support level:

```{r snp-groups}
support_table <- table(Support = dat$Support)
knitr::kable(as.data.frame(support_table),
             caption = "SNP counts by support level")
```

# Credible sets

Each fine-mapping method assigns SNPs to credible sets (CS). A value of
`0` means the SNP is not in any CS; `1` = first CS, `2` = second CS, etc.

```{r credible-sets}
## Count SNPs in each method's credible sets
cs_summary <- data.frame(
    Method = c("ABF", "SUSIE", "FINEMAP"),
    SNPs_in_CS = c(
        sum(dat$ABF.CS > 0, na.rm = TRUE),
        sum(dat$SUSIE.CS > 0, na.rm = TRUE),
        sum(dat$FINEMAP.CS > 0, na.rm = TRUE)
    )
)
knitr::kable(cs_summary, caption = "Number of SNPs in credible sets per method")
```

# Working with multiple loci

`echodata` also bundles LRRK2 and MEX3C loci.
You can combine them to simulate multi-locus analysis:

```{r multi-loci}
bst1 <- data.table::copy(echodata::BST1); bst1[, Locus := "BST1"]
lrrk2 <- data.table::copy(echodata::LRRK2); lrrk2[, Locus := "LRRK2"]
mex3c <- data.table::copy(echodata::MEX3C); mex3c[, Locus := "MEX3C"]
shared_cols <- Reduce(intersect, list(colnames(bst1), colnames(lrrk2), colnames(mex3c)))
all_loci <- rbind(bst1[, ..shared_cols], lrrk2[, ..shared_cols], mex3c[, ..shared_cols])
## Consensus SNPs per locus
consensus_per_locus <- do.call(rbind, lapply(
    split(all_loci, all_loci$Locus),
    function(x) data.frame(
        Locus = x$Locus[1],
        Total_SNPs = nrow(x),
        Consensus_SNPs = sum(x$Consensus_SNP, na.rm = TRUE),
        Min_P = min(x$P, na.rm = TRUE)
    )
))
knitr::kable(consensus_per_locus, row.names = FALSE,
             caption = "Summary across bundled loci")
```

# LD matrix

A pre-computed LD matrix for BST1 is also available:

```{r ld-matrix}
ld <- echodata::BST1_LD_matrix
cat("LD matrix dimensions:", dim(ld), "\n")
cat("SNPs in LD matrix:", ncol(ld), "\n")
## Show a small corner
knitr::kable(round(ld[1:5, 1:5], 3), caption = "LD matrix (top-left corner)")
```

# Session info

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>
