---
title: "Full pipeline vignette" 
author: "Brian M. Schilder"
date: "Most recent update:<br> `r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    theme: spacelab
    highlight: zenburn 
    code_folding: show 
    toc: true 
    toc_float: true
    smooth_scroll: true
    number_sections: false 
    self_contained: true
editor_options: 
  chunk_output_type: inline
vignette: >
  %\VignetteIndexEntry{full_pipeline_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, error = TRUE} 
root.dir <- "~/Desktop"
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = root.dir,
  fig.height = 12,
  fig.width = 10
)  
knitr::opts_knit$set(root.dir = root.dir)
# knitr::opts_chunk$get("root.dir") 

## Build vignettes
## This will generate a /doc folder with the documentation html files.
## Go to GitHub and enable GitHub pages, selecting "branch/doc"
# devtools::build_vignettes(quiet = F, clean=F)
# https://resources.github.com/whitepapers/github-and-rstudio/

# devtools::build()

# devtools::build_manual()
# devtools::build_site()
```

```{r setup, root.dir="~/Desktop"}
library(echolocatoR) 
```

# Full pipeline  

All examples below use data from the Parkinson's disease GWAS by Nalls et al. (2019).  

## Prepare `top_SNPs` data.frame   

* To enable rapid fine-mapping of many loci, you can create a `top_SNPs` data.frame  
which contains the position of the lead/index SNP within each locus.
* `finemap_loci()` (see next step) will then use this info to extract subsets of the   
full GWAS/QTL summary statistics using windows centered on each lead/index SNP.
* The `topSS` argument can either be a data.frame, or a path to a topSS file saved somehwere. Most common tabular data formats (e.g. .tsv, .csv, .xlsx) are accepted.  



```{r  Prepare `top_SNPs` data.frame} 
data("Nalls_top_SNPs");
top_SNPs <- import_topSNPs(
  # topSS = "~/Desktop/Fine_Mapping/Data/GWAS/Nalls23andMe_2019/Nalls2019_TableS2.xlsx",
  topSS = Nalls_top_SNPs,
  chrom_col = "CHR", 
  position_col = "BP", 
  snp_col="SNP",
  pval_col="P, all studies", 
  effect_col="Beta, all studies", 
  gene_col="Nearest Gene", 
  locus_col = "Nearest Gene",
  grouping_vars = c("Locus"),
  remove_variants = "rs34637584") 
head(top_SNPs)
```



## Path to full summary stats file  

* Since a full GWAS summary stats file would be too large to include within *echolocatoR*,  
we instead provide an example subset of the full summary stats.  
* To simulate how you'd actually use your own full summary stats file, we will save our example dataset to your computer (you can change the path to wherever you like).

```{r fullSS}
fullSS_path <- example_fullSS(fullSS_path="~/Desktop/Nalls23andMe_2019.fullSS_subset.tsv") 
```

## Run fine-mapping pipeline  

For a full description of all arguments, see `?finemap_loci`.  

Here are some key arguments:  

* *results_dir*: Where you want to store all of your results.  
* *finemap_methods*: Which fine-mapping methods you want to run (currently includes ABF, FINEMAP, SUSIE, POLYFUN_SUSIE, and COJO).  
* *bp_distance*: Controls window size. Specifically, `bp_distance` is the number of basepairs upstream/downstream you want to extract for each locus. For example, if you want a 2Mb window (+/- 1Mb from the lead/index SNP in `top_SNPs`), set `bp_distance=1e+06`.  
* *plot.zoom*: Zoom in/out from the center of each locus when producing the multiview plot.  
You can adjust this separately from `bp_distance` so that you don't have rerun the whole pipeline each time (locus subsets, LD matrices, and fine-mapping results are all automatically saved in locus-specific folders).  

```{r Run fine-mapping pipeline}

Nalls23andMe_2019.results <- finemap_loci(# GENERAL ARGUMENTS 
                                          top_SNPs = top_SNPs,  
                                          results_dir = "~/Desktop/results",
                                          loci = c("BST1","MEX3C"),# top_SNPs$Locus,
                                          dataset_name = "Nalls23andMe_2019",
                                          dataset_type = "GWAS",  
                                          force_new_subset = F,
                                          force_new_LD = F,
                                          force_new_finemap = F,
                                          remove_tmps = F,
                                          
                 # SUMMARY STATS ARGUMENTS
                 fullSS_path = fullSS_path,
                 query_by ="tabix",
                 chrom_col = "CHR", position_col = "POS", snp_col = "RSID",
                 pval_col = "p", effect_col = "beta", stderr_col = "se",
                 freq_col = "freq", MAF_col = "calculate",
                 A1_col = "A1",
                 A2_col = "A2",
                 
                 # FILTERING ARGUMENTS
                 bp_distance = 500000*2,
                 min_MAF = 0.001,  
                 trim_gene_limits = F,
                 
                 # FINE-MAPPING ARGUMENTS
                 ## General
                 finemap_methods = c("ABF","FINEMAP","SUSIE","POLYFUN_SUSIE"), 
                 n_causal = 5,
                 PP_threshold = .95,
                 ## tool-specific
                 finemap_args=list("FINEMAP"=list("--n-threads"=4),
                                   "SUSIE"=list("scaled_prior_variance"=.001, 
                                                "max_iter"=100)),
                 
                 # LD ARGUMENTS 
                 LD_reference = "UKB",#"1KGphase1",
                 superpopulation = "EUR",
                 download_method = "axel",
                 
                 # PLOT ARGUMENTS 
                 ## general   
                 plot.types=c("fancy"),
                 ## Generate multiple plots of different window sizes; 
                 ### all SNPs, 4x zoomed-in, and a 50000bp window
                 plot.zoom = c("all","4x","10x"),
                 ## XGR
                 # plot.XGR_libnames=c("ENCODE_TFBS_ClusteredV3_CellTypes"), 
                 ## Roadmap
                 plot.Roadmap = F,
                 plot.Roadmap_query = NULL,
                 # Nott et al. (2019)
                 plot.Nott_epigenome = T, 
                 plot.Nott_show_placseq = T, 
                 
                 verbose = F
                 )
```

# Session info  

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
sessioninfo::session_info()
```

</details>


