---
title: "Plotting fine-mapping results"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Plotting fine-mapping results}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include=TRUE}
library(echolocatoR)

can_plot <- tryCatch({
    requireNamespace("echoplot", quietly = TRUE) &&
    requireNamespace("ggplot2", quietly = TRUE) &&
    requireNamespace("patchwork", quietly = TRUE)
}, error = function(e) FALSE)
```

# Overview

`echoplot` provides multi-track locus plots that combine GWAS
association signals, fine-mapping posterior probabilities, LD structure,
and gene models in a single figure. This vignette demonstrates common
plotting workflows using bundled example data — **no internet
connection is required**.

For the full fine-mapping pipeline, see `vignette("echolocatoR")`.
For interpreting results, see `vignette("explore_results")`.

# Load example data

```{r load-data}
dat <- echodata::BST1
LD_matrix <- echodata::BST1_LD_matrix
locus_dir <- file.path(tempdir(), echodata::locus_dir)
```

# Basic locus plot

The main function `echoplot::plot_locus()` creates a multi-panel figure
with GWAS p-values, fine-mapping posterior probabilities, and gene tracks.
SNPs are colored by LD r² with the lead SNP.

```{r basic-plot, fig.height=8, fig.width=7, eval=can_plot}
plt <- tryCatch(
    echoplot::plot_locus(
        dat = dat,
        locus_dir = locus_dir,
        LD_matrix = LD_matrix,
        show_plot = FALSE,
        save_plot = FALSE,
        verbose = FALSE
    ),
    error = function(e) { message("plot_locus error: ", e$message); NULL }
)
if (!is.null(plt)) plt[["1x"]]
```

# Customizing the plot

## Multiple zoom levels

You can generate plots at different zoom levels to focus on the
fine-mapped region:

```{r zoom, fig.height=8, fig.width=7, eval=can_plot}
plt_zoom <- tryCatch(
    echoplot::plot_locus(
        dat = dat,
        locus_dir = locus_dir,
        LD_matrix = LD_matrix,
        zoom = c("1x", "4x"),
        show_plot = FALSE,
        save_plot = FALSE,
        verbose = FALSE
    ),
    error = function(e) { message("plot_locus error: ", e$message); NULL }
)
if (!is.null(plt_zoom)) plt_zoom[["4x"]]
```

## Without LD coloring

If you don't have an LD matrix, the plot still works — SNPs are
shown without r² coloring:

```{r no-ld, fig.height=8, fig.width=7, eval=can_plot}
plt_nold <- tryCatch(
    echoplot::plot_locus(
        dat = dat,
        locus_dir = locus_dir,
        LD_matrix = NULL,
        show_plot = FALSE,
        save_plot = FALSE,
        verbose = FALSE
    ),
    error = function(e) { message("plot_locus error: ", e$message); NULL }
)
if (!is.null(plt_nold)) plt_nold[["1x"]]
```

# Saving plots

To save a plot to disk:

```{r save-example, eval=FALSE}
plt <- echoplot::plot_locus(
    dat = dat,
    locus_dir = locus_dir,
    LD_matrix = LD_matrix,
    save_plot = TRUE,
    plot_format = "png",
    dpi = 300,
    height = 12,
    width = 10,
    show_plot = FALSE
)
```

# Next steps

- Fine-map your own loci: `vignette("echolocatoR")`
- Explore results in detail: `vignette("explore_results")`
- Summarise across loci: `vignette("summarise")`
- Learn about sub-packages: `vignette("echoverse_modules")`

# Session info

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>
