---
title: "Plot locus"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>" 
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Plot locus} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=TRUE}
library(echolocatoR) 
```

# Plotting loci with *echolocatoR*

*echolocatoR* contains various functions that can be used separately  
from the comprehensive `finemap_loci()` pipeline.  

Generate a multi-view plot of a given locus using `echoplot::plot_locus()`.  

* You can mix and match different tracks and annotations using the different arguments 
(see `?echoplot::plot_locus` for details).  

The plot is centered on the lead/index SNP. If a list is supplied to plot.xoom
* `echoplot::plot_locus()` returns a series of `ggplot` objects bound together with [`patchwork`](https://patchwork.data-imaginist.com). One can further modify this object using `ggplot2` functions like `+ theme()`.
  + The modifications will be applied to all tracks at once.  
  
* Save a high-resolution versions the plot by setting `save_plot=T`.  
  + Further increase resolution by adjusting the `dpi` argument (*default=300*).
  + Files are saved in *jpg* format by default, but users can specify their preferred file format (e.g. `file_format="png"`)
  + Adjust the `height` and `width` of the saved plot using these respective arguments.
  + The plot will be automatically saved in the locus-specific directory as:  
  *multiview_<locus>_<plot.zoom>.jpg*.
   
## Load example data   

Load example dataset of the results from fine-mapping the BST1 locus with `finemap_loci()`.
Original data comes from the recent Nalls et al. (2019) Parkinson's disease GWAS 
(see `?BST1` for details).

```{r Load example data}
dat <- echodata::BST1
LD_matrix <- echodata::BST1_LD_matrix
locus_dir <- file.path(tempdir(), echodata::locus_dir)
LD_reference <- "UKB" # Used for naming saved plots 
```


## Full window

```{r trk_plot}
trk_plot <- echoplot::plot_locus(dat = dat, 
                                 LD_matrix = LD_matrix, 
                                 LD_reference = LD_reference,
                                 locus_dir = locus_dir,
                                 zoom = "10x") 
```

## At multiple zooms

* You can easily generate the same locus plot at multiple zoomed in views by supplying a list to `plot.zoom`.  
* This list can be composed of zoom multipliers (e.g. `c("1x", "2x")`), window widths in units of basepairs (e.g. `c(5000, 1500)`), or a mixture of both (e.g. `c("1x","4x", 5000, 2000)`). 
* Each zoom view will be saved individually with its respective scale as the suffix (e.g. `multiview.BST1.UKB.4x.jpg`).  
* Each zoom view is stored as a named item within the returned list.  

```{r trk_plot_zoom, eval=FALSE}
trk_plot_zooms <- echoplot::plot_locus(dat = dat, 
                                       LD_matrix = LD_matrix, 
                                       LD_reference = LD_reference,
                                       locus_dir = locus_dir,
                                       zoom = c("1x","5x","10x")) 
names(trk_plot_zooms) # Get zoom view names
```

## Return as list  

* For even further control over each track of the multi-view plot, specify `echoplot::plot_locus(..., return_list=T)` to instead return a named list (nested within each zoom view list item) of `ggplot` objects which can each be modified individually. 
* Once you've made your modifications, you can then bind this list of plots back together with `patchwork::wrap_plots(tracks_list, ncol = 1)`.  

```{r trk_plot_list}
trk_plot_list <- echoplot::plot_locus(dat = dat, 
                                      LD_matrix = LD_matrix, 
                                      LD_reference = LD_reference,
                                      locus_dir = locus_dir,
                                      zoom = "10x",
                                      show_plot = FALSE,
                                      return_list = TRUE)  
```

```{r extract view}
view1_list <- trk_plot_list[["10x"]]
names(view1_list) # Get track names from a particular zoom view
```

Modify a specific tracks within a view. 

```{r modify track} 
library(ggplot2)
library(patchwork)
# Modify your selected track
modified_track <- view1_list$GWAS + 
    ggplot2::labs(title = "Modified GWAS") +  
    ggplot2::scale_color_gradient2(low = "purple", 
                                   mid = "magenta",
                                   high = "white",
                                   midpoint = .5) +
    ggplot2::theme(title = ggplot2::element_text(hjust = .5),
                   panel.background = ggplot2::element_rect(fill = "black"))
# Put it back into your track list
view1_list[["GWAS"]] <- modified_track
# Remove a plot you don't want
view1_list[["Genes"]] <- NULL
# Specify the relative heights of each track (make sure it matches your new # of plots!)
track_heights <- c(.3,.1,.3,1)
# Bind them back together and plot
fused_plot <- patchwork::wrap_plots(view1_list, 
                                    heights = track_heights,
                                    ncol = 1)
print(fused_plot)
```


## Using XGR annotations   

* Whenever you use annotation arguments (e.g. `xgr_libnames`,`roadmap`,`nott_epigenome`)
the annotations that overlap with your locus will automatically be saved as `GRanges` objects in a locus-specific subdirectory:  
*results/<dataset_type>/<dataset_name>/<locus>/annotation* 
* If a selected annotation has previously been downloaded and stored for that locus, `echoplot::plot_locus()` will automatically detect and import it to save time.   

```{r trk_plot.xgr}
trk_plot.xgr <- echoplot::plot_locus(dat = dat, 
                                     LD_matrix = LD_matrix, 
                                     LD_reference = LD_reference,
                                     locus_dir = locus_dir,  
                                     xgr_libnames = c("ENCODE_TFBS_ClusteredV3_CellTypes"), 
                                     zoom = "10x")
```

## Using [Roadmap](http://www.roadmapepigenomics.org/data/tables/brain) annotations

* Using the `Roadmap=T` and `Roadmap_query="<query>"` arguments searches the Roadmap for chromatin mark data across various cell-types, cell-lines and tissues.  
* Note that Roadmap queries requires `tabix` to be installed on your machine, or within a conda environment (`conda_env = "echoR"`). 
* Parallelizing these queries across multiple threads speeds up this process (`nThread=<n_cores_available>`), as does reusing previously stored data which is automatically saved to the locus-specific subfolder (`<dataset_type>/<dataset_name>/<locus>/annotations/Roadmap.ChromatinMarks_CellTypes.RDS`).

```{r trk_plot.roadmap, eval=FALSE} 
trk_plot.roadmap <- echoplot::plot_locus(dat = dat, 
                                         LD_matrix = LD_matrix, 
                                         LD_reference = LD_reference,
                                         locus_dir = locus_dir,
                                         zoom = "5x",
                                         
                                         roadmap = TRUE, 
                                         roadmap_query = "monocyte")
```
 
## Using [Nott_2019](https://science.sciencemag.org/content/366/6469/1134.abstract)   annotations   

* Query and plot brain cell type-specific epigenomic assays from 
[Nott et al. (Science, 2019)](https://science.sciencemag.org/content/366/6469/1134.abstract)  
(see `?NOTT_2019.bigwig_metadata` for details).

```{r trk_plot.nott_2019, eval=FALSE} 
trk_plot.nott_2019 <- echoplot::plot_locus(dat = dat,  
                                 LD_matrix = LD_matrix, 
                                 LD_reference = LD_reference,
                                 locus_dir = locus_dir, 
                                 zoom = "10x",
                                 
                                 nott_epigenome = TRUE,  
                                 nott_binwidth = 200,
                                 nott_regulatory_rects = TRUE, 
                                 nott_show_placseq = TRUE) 
```

## Using QTL datasets  

* Plot multiple QTL p-value columns (or really P-value columns from any kind of datset).  
* Each QTL dataset will be plotted as a new track.

```{r trk_plot.QTL}
dat1 <- data.table::copy(dat)
dat2 <- data.table::copy(dat)
# Make fake QTL P-values for the sake a demonstration
dat1$P <- abs(jitter(dat1$P, amount = 1e-15))
dat2$P <- abs(jitter(dat2$P, amount = 1e-16))
dat_list <- list("fake_eQTL"=dat1, 
                 "fake_sQTL"=dat2)

trk_plot.qtl <- echoplot::plot_locus_multi(dat_ls = dat_list, 
                                           LD_ls = list(LD_matrix,LD_matrix),  
                                           locus_dir = locus_dir,
                                           zoom = "10x")
```

<hr>

# Session info  

<details>  

```{r Session Info, attr.output='style="max-height: 400px;"'}
utils::sessionInfo()
```

<details>

<br>
