---
title: "Plotting vignette"
author: "Brian M. Schilder"
date: "Most recent update:<br> `r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    theme: spacelab
    highlight: zenburn 
    code_folding: show 
    toc: true 
    toc_float: true
    smooth_scroll: true
    number_sections: false 
    self_contained: true
editor_options: 
  chunk_output_type: inline
vignette: >
  %\VignetteIndexEntry{plotting_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, error = TRUE, dpi=300}
root.dir <- "~/Desktop"
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = root.dir, 
  fig.height = 12,
  fig.width = 10
)  
knitr::opts_knit$set(
  root.dir = root.dir
  )
# knitr::opts_chunk$get("root.dir")
# devtools::build_vignettes(quiet = F)
```

```{r setup, root.dir="~/Desktop", dpi=300}
library(echolocatoR) 
```

# Plotting loci with *echolocatoR*

*echolocatoR* contains various functions that can be used separately  
from the comprehensive `finemap_loci()` pipeline.  

Generate a multi-view plot of a given locus using `PLOT.locus()`.  

* You can mix and match different tracks and annotations using the different arguments 
(see `?PLOT.locus` for details).  

The plot is centered on the lead/index SNP. If a list is supplied to plot.xoom
* `PLOT.locus()` returns a series of `ggplot` objects bound together with [`patchwork`](https://patchwork.data-imaginist.com). One can further modify this object using `ggplot2` functions like `+ theme()`.
  + The modifications will be applied to all tracks at once.  
  
* Save a high-resolution versions the plot by setting `save_plot=T`.  
  + Further increase resolution by adjusting the `dpi` argument (*default=300*).
  + Files are saved in *jpg* format by default, but users can specify their preferred file format (e.g. `file_format="png"`)
  + Adjust the `height` and `width` of the saved plot using these respective arguments.
  + The plot will be automatically saved in the locus-specific directory as:  
  *multiview_<locus>_<plot.zoom>.jpg*.
   
## Load example data   

Load example dataset of the results from fine-mapping the BST1 locus with `finemap_loci()`.
Original data comes from the recent Nalls et al. (2019) Parkinson's disease GWAS (see `?BST1` for details).

```{r Load example data   }
library(ggplot2)
data("BST1"); data("LD_matrix"); data("locus_dir");
locus_dir <- file.path("~/Desktop",locus_dir)
finemap_DT <- BST1  
LD_reference <- "UKB" # Used for naming saved plots
plot.zoom = "10x"
show_plot <- F


# locus <- "DNAH17"
# locus_dir <- file.path("~/Desktop/Fine_Mapping/Data/GWAS/Nalls23andMe_2019",locus)
# LD_matrix <- readRDS(file.path(locus_dir,"plink/UKB_LD.RDS"))
# finemap_DT <- data.table::fread(file.path(locus_dir,"Multi-finemap/Multi-finemap_results.txt"))
# finemap_DT <- find_consensus_SNPs(finemap_DT) 
# finemap_DT <- update_CS_cols(finemap_DT)
```


## Full window

```{r trk_plot, attr.output='style="max-height: 200px;"'}
trk_plot <- PLOT.locus(finemap_dat=finemap_DT, 
                       LD_matrix=LD_matrix, 
                       LD_reference=LD_reference,
                       locus_dir=locus_dir, 
                       XGR_libnames=NULL,  
                       save_plot=F,
                       show_plot=show_plot,
                       plot.zoom=plot.zoom) 
```

```{r}
print(trk_plot)
```

## At multiple zooms

* You can easily generate the same locus plot at multiple zoomed in views by supplying a list to `plot.zoom`.  
* This list can be composed of zoom multipliers (e.g. `c("1x", "2x")`), window widths in units of basepairs (e.g. `c(5000, 1500)`), or a mixture of both (e.g. `c("1x","4x", 5000, 2000)`). 
* Each zoom view will be saved individually with its respective scale as the suffix (e.g. `multiview.BST1.UKB.4x.jpg`).  
* Each zoom view is stored as a named item within the returned list.  

```{r trk_plot_zoom, attr.output='style="max-height: 200px;"'}
trk_plot_zooms <- PLOT.locus(finemap_dat=finemap_DT, 
                             LD_matrix=LD_matrix, 
                             LD_reference=LD_reference,
                             locus_dir=locus_dir, 
                             XGR_libnames=NULL, 
                             save_plot=F,
                             show_plot=show_plot,
                             plot.zoom = c("1x","5x","10x")) 
names(trk_plot_zooms) # Get zoom view names
```

```{r}
print(trk_plot_zooms)
```


## Return as list  

* For even further control over each track of the multi-view plot, specify `PLOT.locus(..., return_list=T)` to instead return a named list (nested within each zoom view list item) of `ggplot` objects which can each be modified individually. 
* Once you've made your modifications, you can then bind this list of plots back together with `patchwork::wrap_plots(tracks_list, ncol = 1)`.  

```{r trk_plot_list, attr.output='style="max-height: 200px;"'}
trk_plot_list <- PLOT.locus(finemap_dat=finemap_DT, 
                             LD_matrix=LD_matrix, 
                             LD_reference=LD_reference,
                             locus_dir=locus_dir, 
                             XGR_libnames=NULL, 
                             save_plot=F,
                             show_plot=show_plot,
                             plot.zoom=plot.zoom,
                             return_list=T)  
```

```{r extract view}
view1_list <- trk_plot_list[[plot.zoom]]
names(view1_list) # Get track names from a particular zoom view
```

Modify a specific tracks within a view. 

```{r modify track} 
# Modify your selected track
modified_track <- view1_list$GWAS + 
                      labs(title = "Modified GWAS") + 
                      theme_dark() +
                      theme(title = element_text(hjust = .5))
# Put it back into your track list
view1_list[["GWAS"]] <- modified_track
# Remove a plot you don't want
view1_list[["Genes"]] <- NULL
# Specify the relative heights of each track (make sure it matches your new # of plots!)
track_heights <- c(.3,.1,.3,1)

# Bind them together and plot
fused_plot <- patchwork::wrap_plots(view1_list, 
                                    heights = track_heights,
                                    ncol = 1)
print(fused_plot)
```


## Using XGR annotations   

* Whenever you use annotation arguments (e.g. `XGR_libnames`,`Roadmap`,`Nott_epigenome`)
the annotations that overlap with your locus will automatically be saved as `GRanges` objects in a locus-specific subdirectory:  
*results/<dataset_type>/<dataset_name>/<locus>/annotation* 
* If a selected annotation has previously been downloaded and stored for that locus, `PLOT.locus()` will automatically detect and import it to save time.   

```{r trk_plot.xgr, attr.output='style="max-height: 200px;"'}
trk_plot.xgr <- PLOT.locus(finemap_dat=finemap_DT, 
                           LD_matrix=LD_matrix, 
                           LD_reference=LD_reference,
                           locus_dir=locus_dir, 
                           XGR_libnames=c("ENCODE_TFBS_ClusteredV3_CellTypes"), 
                           save_plot=F,
                           show_plot=show_plot,
                           plot.zoom=plot.zoom)
```

```{r}
print(trk_plot.xgr)
```

## Using [Roadmap](http://www.roadmapepigenomics.org/data/tables/brain) annotations

```{r trk_plot.roadmap, attr.output='style="max-height: 200px;"'} 
trk_plot.roadmap <- PLOT.locus(finemap_dat=finemap_DT, 
                               LD_matrix=LD_matrix, 
                               LD_reference=LD_reference,
                               locus_dir=locus_dir, 
                               XGR_libnames=NULL, 
                               Roadmap=T, 
                               Roadmap_query="monocyte", 
                               save_plot=F, 
                               show_plot=show_plot,
                               plot.zoom=plot.zoom)
```

```{r}
print(trk_plot.roadmap)
```

## Using [Nott_2019](https://science.sciencemag.org/content/366/6469/1134.abstract)   annotations   

* Query and plot brain cell type-specific epigenomic assays from 
[Nott et al. (Science, 2019)](https://science.sciencemag.org/content/366/6469/1134.abstract)  
(see `?NOTT_2019.bigwig_metadata` for details).

```{r trk_plot.nott_2019, attr.output='style="max-height: 200px;"'} 
trk_plot.nott_2019 <- PLOT.locus(finemap_dat=finemap_DT,  
                                 LD_matrix=LD_matrix, 
                                 LD_reference=LD_reference,
                                 locus_dir=locus_dir,
                                 XGR_libnames=NULL, 
                                 Nott_epigenome=T,  
                                 Nott_binwidth = 100,
                                 Nott_regulatory_rects = T, 
                                 Nott_show_placseq = T,
                                 save_plot=T,
                                 show_plot=show_plot,
                                 plot.zoom=plot.zoom)
 
```

```{r}
print(trk_plot.nott_2019)
```

## Using QTL datasets  

* Plot multiple QTL p-value columns (or really P-value columns from any kind of datset).  
* Each QTL dataset will be plotted as a new track.

```{r trk_plot.QTL, attr.output='style="max-height: 200px;"'}
# Make fake QTL P-values for the sake a demonstration
finemap_DT$fake_eQTL.P <- finemap_DT$P  * c(1,.9,.7)
finemap_DT$fake_sQTL.P <- finemap_DT$P  * c(1,.8,.5)

trk_plot.qtl <- PLOT.locus(finemap_dat=finemap_DT, 
                           LD_matrix=LD_matrix, 
                           LD_reference=LD_reference,
                           locus_dir=locus_dir,
                           QTL_prefixes=c("fake_eQTL.","fake_sQTL."),
                           XGR_libnames=NULL, 
                           save_plot=F,
                           show_plot=show_plot,
                           plot.zoom=plot.zoom)
```

```{r}
print(trk_plot.qtl)
```

<details>
<summary><h1>Session Info</h1></summary>  

```{r Session Info, attr.output='style="max-height: 200px;"'}
sessioninfo::session_info()
```
<details>

