---
title: "QTLs"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{QTLs}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=TRUE}
library(echolocatoR)

has_internet <- tryCatch(
    !is.null(curl::nslookup("github.com", error = FALSE)),
    error = function(e) FALSE
)
```


# QTL pipeline

- Here, we will use GWAS-eQTL colocalization results provided via the [*echolocatoR Fine-mapping Portal*](https://github.com/RajLabMSSM/Fine_Mapping_Shiny),
from the publication:

> Lopes, K.d.P., Snijders, G.J.L., Humphrey, J. et al. Genetic analysis of the human microglial transcriptome across brain regions, aging and disease pathologies. Nat Genet 54, 4–17 (2022). [https://doi.org/10.1038/s41588-021-00976-y](https://doi.org/10.1038/s41588-021-00976-y)

## Import QTL data

This data is actually merged GWAS-QTL colocalization results,
but it contains all of the necessary columns from the original eQTL
summary stats that we need to perform eQTL fine-mapping.

```{r import-qtl, eval=has_internet}
coloc_res <- echodata::get_Kunkle2019_coloc(return_path = TRUE)
```

## Prepare `colmap`

Prepare a column mapping object for the summary statistics.
We'll reuse this for both the `import_topSNPs` and `finemap_loci` steps.

```{r prepare-colmap, eval=has_internet}
colmap <- echodata::construct_colmap(
      CHR = "chr",
      POS = "pos",
      N = "qtl.N",
      SNP = "snp",
      P = "qtl.pvalues",
      Effect = "qtl.beta",
      StdErr = "qtl.varbeta",
      MAF = "qtl.MAF",
      Locus = "Locus",
      Gene = "gene")
```


## Prepare `top_SNPs` data.frame

- In this case, we don't have a top SNPs file ready.
So we're just going to make one directly from the full summary stats file itself
(*NOTE*: You can only use this approach if you can fit the entire file in memory).
- In this case, you'll want to make sure to set `grouping_vars=c("Locus","Gene")` so that you get top SNPs for each eGene-locus pair (not just one SNP per locus).

```{r Prepare top_SNPs data.frame, eval=has_internet}
topSNPs <- echodata::import_topSNPs(
  topSS = coloc_res$path,
  colmap = colmap,
  ## Important for QTLs: group by both Locus and Gene
  grouping_vars = c("Locus","Gene"))
head(topSNPs)
```

## Run fine-mapping pipeline

The following call runs the full fine-mapping pipeline on QTL data.
This is computationally expensive and may take **10–30 minutes per locus**.

```{r finemap-qtl, eval=FALSE}
res <- echolocatoR::finemap_loci(fullSS_path = coloc_res$path,
                                 topSNPs = topSNPs,
                                 ## Let's just fine-map 1 locus for demo purposes
                                 loci = topSNPs$Locus[1],
                                 dataset_name = "Kunkle_2019.microgliaQTL",
                                 dataset_type = "QTL",
                                 bp_distance = 1000,
                                 colmap = colmap,
                                 show_plot = TRUE,
                                 finemap_methods = c("ABF","FINEMAP","SUSIE") )
```


# Session info

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>
