---
title: "Summarise"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>" 
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Summarise} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = TRUE} 
library(echolocatoR)  
```

# Download data  

# Summarise

Using pre-merged data for vignette speed. 

```{r}
merged_DT <- echodata::get_Nalls2019_merged()
```

## `get_SNPgroup_counts()`

Get the number of SNPs for each SNP group per locus.  
It also prints the mean number of SNPs for each SNP group across all loci.  
**NOTE**: You will need to make sure to set `merge_finemapping_results(minimum_support=1)`
in the above step to get accurate counts for all SNP groups.  

```{r get_SNPgroup_counts()}
snp_groups <- echodata::get_SNPgroup_counts(merged_DT = merged_DT)
```

## `get_CS_counts()` 

County the number of tool-specific and UCS Credible Set SNPs per locus.  

```{r get_CS_counts()}
UCS_counts <- echodata::get_CS_counts(merged_DT = merged_DT)
knitr::kable(UCS_counts)
```


# Plot  

- The following functions each return a list containing both the `...$plot` and the `...$data`
used to make the plot.  
- Where available, `snp_filter` allows user to use any filtering argument (supplied as a string)
to subset the data they want to use in the plot/data.  


## `SUMMARISE.coloc_nominated_eGenes()`  

If you ran colocalization tests with `echolocatoR` (via `catalogueR`) you can use those results to 
come up with a top QTL nominated gene for each locus (potentially implicating that gene in your phenotype).  

### Download COLOC results from [echolocatoR Fine-mapping Portal](https://rajlab.shinyapps.io/Fine_Mapping_Shiny).    

```{r}
coloc_res <- echodata::get_Nalls2019_coloc() 
```


## Multi-plot  
 

```{r super_summary_plot(), fig.height=12,, fig.width=7}
super_plot <- echoannot::super_summary_plot(merged_DT = merged_DT, 
                                            coloc_results = coloc_res,
                                            plot_missense = FALSE)
```


# Session info 

<details>

```{r Session Info, attr.output='style="max-height: 200px;"'}
utils::sessionInfo()
```

</details>

